# Caret Generic Workflow  
Author:    Neal Cariello  
Purpose:   Develop Generic Workflow For Caret Package To Examine Multiple ML Methods For Classification Datasets  
Code path on Cariello machine: F:\\Google Drive\\R\\caret\\Generic Caret Workflow  
Date: 2015_07_02

This R Markdown code is based "Predictive Modeling with R and the caret package" in useR! 2013 by Max Kuhn


## General caret info
Package caret: http://topepo.github.io/caret/index.html  
R taskview on machine learning: http://cran.r-project.org/web/views/MachineLearning.html  

## R and caret versions
version 3.2.0 (2015-04-16) "Full of Ingredients"  
Platform: x86_64-w64-mingw32/x64 (64-bit)  
Running under: Windows 7 x64 (build 7601) Service Pack 1  
caret version 6.0.47

mtry is a tuning parameter mostly for Random Forest. 
mtry the number of variables randomly sampled as candidates at each split.  
For complete list of methods using mtry, see  http://topepo.github.io/caret/modelList.html

## Set Up Data And TrainContol Method
See http://topepo.github.io/caret/training.html  

## Load data, set up for parallel processing and examine variance and correlations between variables for possible variable deletion
### Using segmentationData included in caret package

```{r, Get Data and Delete Cols}

# clear the decks.
rm(list=ls(all=TRUE))  

USE_SEGMENTATION_DATA_IN_CARET <- TRUE
#  USE_SEGMENTATION_DATA_IN_CARET <- FALSE

set.seed(123)

require(caret)
require(ggplot2)
require(corrplot)     # for examination of variable correlations
require(doParallel)   # for parallel computing

# Use multi-core support
registerDoParallel()
getDoParWorkers()

sessionInfo()
packageVersion("caret")

# segmentationData dataset is provided in caret
data(segmentationData)
dataRaw <- segmentationData

if (USE_SEGMENTATION_DATA_IN_CARET) {
  # get rid of the cell identifier as suggested by Kuhn
  dataRaw$Cell <- NULL
}

dim(dataRaw)
head(dataRaw)
summary(dataRaw)
str(dataRaw)

# Check for variables with near zero variance
# Data needs to be all numeric 
temp <- subset(dataRaw, select = -c(Case, Class))

nzv <- nearZeroVar(temp)  # returns number of near zero variance cols
if( length(nzv) == 0) {nzv <- 0}
print(paste("Number of columns with near zero variance is: ", nzv))

# Examine highly correlated variables for possible exclusion
# Again, data needs to be all numeric

correlations <- cor(temp)
corrplot(correlations, 
         order="hclust",
         mar = c(3, 4, 1, 3),
         title=paste("Correlations of", 
                     length(names(temp)), "variables in dataset"),
          tl.cex = 0.6,     # labels text size
         cex.main=1         # title text size
)

# Change cutoff as you see fit
highlyCorrelated <- findCorrelation(correlations, cutoff=0.9)
print(paste("Highly correlated variables are:",   
            names(temp)[highlyCorrelated])
)

# If columns are to be removed, it must be done in dataRaw dataframe
print("Delete columns here")

# IMPORTANT - df for next code segment must be "dataCleanRemoveCols"
dataCleanRemoveCols <- dataRaw

```

## Scale, transform and impute data

This may not be necessary for your data  
dataCleanRemoveCols is dataframe required for this operation  

```{r, scale and transform data}

# Run this chunk ?
RUN_SCALING <- FALSE
# RUN_SCALING <- TRUE

# dataframe dataCleanRemoveCols is needed here

# See caret documentation for up-to-data documentation for methods.

# Methods available now are "Box-Cox", "YeoJohnson", "expoTrans", "center", "scale", "range", "knnImpute", "bag-Impute",
# "medianImpute", "pca", "ica" and "spatialSign"

# Note imputation methods above


#  >>>>>>>>>>>>>>> IMPORTANT - df must be all numeric for scaling, so you may need to modify your data  <<<<<<<<<<<<<<
# centering and scaling are pretty standard. modify as you see fit

if(RUN_SCALING) {
  dataPreProc <- preProcess(dataCleanRemoveCols, method = c("center", "scale"))
  print("Scaling applied")
}

if(!RUN_SCALING) {
  dataPreProc <- dataCleanRemoveCols
  print("No scaling applied")
}
# dataframe name to next chunk must be dataPreProc


```

##  Create training and test datasets
### Dataframe must be named dataCleanRemoveCols

```{r, createDataPartion}

# dataframe in for this chunk must be named dataCleanRemoveCols

# NOTE that the segmentation data in the caret package already has the training and test sets defined, presumably for
#   reproducibilty.  Training and test are in the variable "Case". Delete this column and proceed with the 
#   creation of the Training and Test sets

#  USING_CARET_SEGMENTATION_DATA <- FALSE
USING_CARET_SEGMENTATION_DATA <- TRUE

if (USING_CARET_SEGMENTATION_DATA) { dataCleanRemoveCols$Case <- NULL }

# For caret example, "Class" is what we are trying to predict

inTrain <- createDataPartition(dataCleanRemoveCols$Class,
                p = 0.75,
                list=FALSE)  # output is set of integers for the rows now in the training set

trainingData <- dataCleanRemoveCols[  inTrain ,]
testingData  <- dataCleanRemoveCols[ -inTrain ,]  

trainingClass <- trainingData$Class
testingClass  <- testingData$Class

```


## Setup trainControl  
As of June 30 2015, the following methods exist for TrainControl  
"boot", "cv", "LOOCV", "LGOCV", "repeatedcv", "timeslice", "none" and "oob"

```{r, train_control}

trn_ctl <- trainControl(method          = "repeatedcv",
                        repeats         = 5,
                        number          = 10,
                        classProbs      = TRUE,               # necessary to get confusion matrix
                        summaryFunction = twoClassSummary     # necessary to get confusion matrix
                        )   
                

```

## Run models using train()
This function sets up a grid of tuning parameters for a number of classification and regression  
routines, fits each model and calculates a resampling based performance measure.  

### Model RPART: Recursive Partitioning and Regression Trees  

```{r, run model rpart}

TUNE_LEN <- 10

startTime <- Sys.time()

rpartFit <- train(Class ~ . , 
                  method     = "rpart", 
                  data       = trainingData,
                  # metric     = "ROC",
                  tuneLength = TUNE_LEN,
	                trControl  = trn_ctl
                  )

endTime <- Sys.time()
print("Time to run rpart")
endTime - startTime

```

## K-Nearest Neighbor

```{r, knn}

startTime <- Sys.time()

knnFit <- train(Class ~ . , 
                  method     = "knn", 
                  data       = trainingData,
                  # metric     = "ROC",
                  tuneLength = TUNE_LEN,
                  trControl  = trn_ctl
                  )

endTime <- Sys.time()
print("Time to run knn")
endTime - startTime


```

## Support Vector Machine

```{r, svm}

svmFit <- train(type ~., 
  method = "svmRadial", 
  data = DATA_IN,
	tuneLength = 5,
	trControl = trainControl(
		method = TRAIN_CONTROL_METHOD, indexOut = train))

attributes(svmFit)
svmFit
svmFit$finalModel

```

## Artificial Neural Network

```{r, ann}

nnetFit <- train(type ~ ., 
  method = "nnet", 
  data = DATA_IN,
	tuneLength = 5,
	trControl = trainControl(
		method = TRAIN_CONTROL_METHOD, indexOut = train))

attributes(nnetFit)
nnetFit
nnetFit$finalModel
plot(nnetFit)

```

## Random Forest

```{r, rf}

randomForestFit <- train(type ~ ., 
  method = "rf", 
  data = DATA_IN,
	tuneLength = 5,
  metric = "ROC",
	trControl = trainControl(
		method = TRAIN_CONTROL_METHOD, 
      indexOut = train,
       classProbs = TRUE))

attributes(randomForestFit)
randomForestFit
randomForestFit$finalModel
plot(randomForestFit)

``` 

## Compare the models

```{r, Compare Models}

compareModels <- resamples(list(
  RPART = rpartFit,
  KNN = knnFit),
  decreasing = TRUE
  )


resamps <- resamples(list(
  ctree=ctreeFit,
  C45=C45Fit,
  SVM=svmFit,
  KNN=knnFit,
  NeuralNet=nnetFit,
  randomForest=randomForestFit),
  decreasing=TRUE)

attributes(resamps)
resamps
summary(resamps)

difs <- diff(resamps)
difs
summary(difs)


```
